---
title: NACHO Data Report: bulk $\delta$^13^C and $\delta$^15^N
output: pdf_document 
geometry: left = 1in,right = 1in,top = 1in,bottom = 1in 
subtitle: XXX
description: XXX
params:
    analysis.dates:
    data.files:
    standard.plots:
    standard.coefficients:
    known.standard.values:
    calibration.coefficients:
    error.analysis.results:
    zeros.CN:
    blanks.CN:    
    standard.CN:
    sample.CN:
    blank.CN:
    run.comments:
      
---
# LABORATORY																						
	UW Facility for Compound-Specific Isotope Analysis of Environmental Samples (known informally a NACHO)
	College of the Environment
	University of Washington
	Director is Gordon Holtgrieve, gholt@uw.edu, 206-227-9930
																						
# METHOD																						
	Your solid samples were analyzed for bulk $\delta$^13^C and $\delta$^15^N on a ThermoFinnigan Delta V with a Carlo Erba elemental analyzer in continuous flow mode following the
	general method of Fry et al. 1992. Automated analysis system for coupled d13C and d15N measurements. Analytical Chemistry 64, 288-291.	

# ANALYSIS																						
	Date of Analysis (MM/DD/YY): 	'r analysis.dates'																				
	Original Filename: 'r data.file'																					
	Reduced Standards Filename:	'r paste0("standardCN_",data.file)'
	Reduced Samples Filename:	'r paste0("sampleCN_",data.file)'
	Run type: bulk C and N																				
	Run comments: 'r run.comments'																					
																						
# REFERENCE MATERIALS																						
	All internationally recognized reference material accepted values can be found at the [CIAAW](http://www.ciaaw.org/). Typically we use 
	IsoLab working standards GA1, GA2, and Bristol Bay Sockeye (salmon) on NACHO.  You can find information about these standards on 
	the [IsoLab web page](https://isolab.ess.washington.edu/resources/standards.php#solid). Below are data specific to this run:
```{r, echo=FALSE, results='asis'}
library(tibble)
library(knitr)

knitr::kable(tibble(known.standard.values), format = "simple", caption="Lab reference materials used in this run and their accepted values.")

#calculate means and SD for both raw and corrected GA1, GA2, and Salmon
knitr::kable(XXX, format = "simple", caption="Lab reference material performance summary for this run.")

knitr::kable(tibble(calibration.coefficients), format = "simple", caption="Linear calibration curve coefficents used for this run (y=mx + b).")

knitr::kable(tibble(error.analysis.results), format = "simple", caption="Accuracy and precision of standards for this run.")

```	
													
# ZEROS & BLANKS
  The table below includes measurable blanks and zeros. Blanks are empty tins while zeros are no tin or sample. The table below, if given, 
  contains the data for blanks and zero from this run. No table indicates blanks and zeros were not measurable. Blank correction has
  not been implemented in the script. 
```{r echo=FALSE}
#knitr::kable(blank.CN, format = "simple", caption="Blank data.")
#knitr::kable(zero.CN, format = "simple", caption="Blank data.") 

```
  

# MASS EFFECTS
  Your data were analysed for effects of sample mass on peak area and isotopic ratios. Ideally, there should be a strong linear response of 
  mass C (or N) in the standard on area 44 (or 28) and no effect of peak area on $\delta$^13^C or $\delta$^15^N.
```{r echo=FALSE}

standard.plots[[1]]
standard.plots[[2]]
#knitr::kable(standard.coefficients, format = "simple", caption="Liner model coefficients of mass effects.")           
               
```

#SAMPLE MASS CHECK
Not implemented yet. P


# SAMPLE DATA
```{r echo=FALSE, results='asis'}
library(tibble)
knitr::kable(tibble(sample.CN), format = "simple", caption="Here are the data you have been waiting for.")

```
															
# DATA REDUCTION DETAILS
```{r, echo=FALSE, warning=F, message=F}
library(knitr)
x <- matrix(c("File rendered on:", format(Sys.time(), "%d %b %Y"),
              "R version:", as.character(getRversion()),
              "HEEL package version:", as.character(packageVersion("HEEL"))), ncol=2, byrow=T)
print(knitr::kable(x, row.names = F))
```






